---
title: "Orange County Air Pollution and Household Income"
Author: Stephan Kadonoff
Date: 10/24/2025
format: html
editor: visual
message: false
warning: false
---

#### Author: Stephan Kadonoff                                                         Date:10/24/2025

## Mapping Environmental Hazards in Orange County

## Environmental Justice: Correlation of poverty and exposure to air pollution.

Environmental challenges are not just limited to public land degradation or pollutant impacts on local species. Often times, the human element is overlooked when discussing negative externalities of air pollution. Within this quarto doc, we seek to analyze and better understand how air pollution impacts the millions of residents who call Orange County home. Beyond simply mapping areas of high air pollution, we also will look at which areas within that same geographic area have high levels of low household income

### 1. Preliminary Data Loading

To begin with, we will want to load relevant libraries, as well as read in the ej_screen geodatabase file into our environment.

```{r}
#| message: false
library(tidyverse)
library(sf)
library(here)
library(tmap)

```

```{r}
#| message: false
#| warning: false
# read in the the geodatabase 
ejscreen <- sf::st_read(here::here("data", "ejscreen", "EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb"))
```

### 2. Filtering data for our region of focus

After the initial steps of loading in the data, we will want to start parsing out only the sections which are relevant for our analysis.

```{r}
#| message: false
#| warning: false
# filter to only include the State and County of focus (Santa Ana) and multiple decimal point variable by 100 to see percentage point values

Orange_County <- ejscreen %>%
  filter(ST_ABBREV == "CA", CNTY_NAME %in% c("Orange County"))
Orange_County$LOWINCPCT <-Orange_County$LOWINCPCT * 100
```

### Creating the Maps

With our data now properly filtered, we can create our two maps. The first is a map showing areas with high levels of air pollution across Orange County, CA and the second map looks at zipcodes with low household income.

```{r}
#| message: false
#| warning: false
# Map poor air quality, RSEI_AIR, within Orange County, with a color gradient system focused on different particulate levels.

tm_shape(Orange_County) +
  tm_polygons(fill = "RSEI_AIR",
              border.col = "black",
              palette = "Reds",
              title = "Air Pollution Index",
              breaks= c(0.00, 199.00, 399.00, 599.00, 799.0, 999.00, 1199.00, 2000.00, 3000.00, 5000.00, 10000.00, 15000.00)) + 
  tm_compass(position = c("left", "top"))+ 
  tm_scalebar(position = c("left", "bottom")) + 
  tm_layout(main.title = "Areas in Orange County CA With High Air Pollution")
```

```{r}
#| message: false
#| warning: false
# Map the percentage of low income designated households by zipcode to determine if there is a spatial correlation between income levels and hazardous air levels. 
tm_shape(Orange_County) +
tm_polygons(fill = "LOWINCPCT", 
            border.col = "black",
            breaks= c(0, 20, 40, 60, 80, 100),
            title = "% of HH Low Income") + 
  tm_compass(position = c("left", "top"))+ 
  tm_scalebar(position = c("left", "bottom")) + 
  tm_layout(main.title = "Map of Low Income Households in Orange County CA")
```
