---
title: "Untitled"
format: html
editor: visual
---

```{r}
install.packages("tmap")
library(tidyverse)
library(sf)
library(here)
library(tmap)
```

```{r}
# read in the the geodatabase 
ejscreen <- sf::st_read(here::here("data", "ejscreen", "EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb"))
```

```{r}
# filter to include the State and County of focus (Santa Ana)

Orange_County <- ejscreen %>%
  dplyr::filter(ST_ABBREV == "CA", CNTY_NAME %in% c("Orange County"))
Orange_County$LOWINCPCT <-Orange_County$LOWINCPCT * 100
```

```{r}
# Map Orange County
tm_shape(Orange_County) +
  tm_polygons(col.scale = "PRMP",
              palette = c("yellow", "red", "darkred"),
              tm_legend("Particulate Matter"),
              breaks=c(0.00, 0.25, 0.5, 0.75, 1.0, 2.00, 3.00, 4.00, 5.00)) +
  tm_compass(position = c("left", "top")) +
  tm_scalebar() +
  tm_layout(main.title = "Particulate Matter in Orange County CA")

```

```{r}


tm_shape(Orange_County) +
tm_polygons(fill = "RSEI_AIR", 
            breaks=c(0.00, 199.00, 399.00, 599.00, 799.0, 999.00, 1199.00, 2000.00, 3000.00, 5000.00, 10000.00, 15000.00), 
            fill.scale = tm_scale(values = c("yellow", "darkred"))) + 
  tm_compass(position = c("left", "top"))+ 
  tm_scalebar(position = c("left", "bottom")) + 
  #tm_graticules() +
  tm_layout(main.title = "Relative Potential Human Health Impacts")
```

```{r}
# Map showing particulate matter/low air quality is generally concentrated in the North sector of the County

tm_shape(Orange_County) +
tm_polygons(fill = "RSEI_AIR",
            border.col = "black",
            fill.scale = tm_scale(values = c("yellow", "darkred")),
            fill.legend = tm_legend(title = "Particulate Matter (PPM)"),
            breaks= c(0.00, 199.00, 399.00, 599.00, 799.0, 999.00, 1199.00, 2000.00, 3000.00, 5000.00, 10000.00, 15000.00)) + 
  tm_compass(position = c("left", "top"))+ 
  tm_scalebar(position = c("left", "bottom")) + 
  #tm_gradicule
  tm_layout(main.title = "Relative Potential Human Health Impacts")
```

```{r}
# besure to include gradules
# col, palette, title, and style
# fill. legend to change text
tm_shape(Orange_County) +
tm_polygons(fill = "LOWINCPCT", 
            breaks= c(0, 20, 40, 60, 80, 100)) + 
  tm_compass(position = c("left", "top"))+ 
  tm_scalebar(position = c("left", "bottom")) + 
  tm_layout(main.title = "Relative Potential Human Health Impacts")
```

These two maps were created to show, if true, a spatial correlation between quantity of potentially harmful particulate matter, and percentage of homes deemed low income. What these maps show is that there is a higher quantity of detected particulates, on average, within the communities found within the north and north-west sections of Orange County. This coincides with a general trend where many of those same areas also have higher levels of residents residing within low income households.

```{r}
#tm_fill (pallete, title)
```
